Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-05 WSSSPE3/report/wg_main_user_community.tex
added line breaks
so it looks better on github in non-edit mode
typo
fixed some quotes

Mozilla Science maintains a "'Working Open' Project Guide" (http://mozillascience.github.io/leadership-training). From the introduction:
\begin{quote}
Working openly with contributors enables your
    community to learn how to build and collaborate together. This
    document is a guideline on how to work openly and involve others
    in your projects with Mozilla. We want to help you engage your
    community in a way that encourages contributors and builds other
    leaders.
  \end{quote}

Several books have been written about software communities:
\begin{itemize}
\item "Art of Community" by Jono Bacon. We could consider distilling this for scientific software.
\item Iain Larmour, from EPSRC inthe UK [[https://www.epsrc.ac.uk/][EPSRC]] (not sure who from mentioned UK Collaborative Computational Projects ([[http://ccp.ac.uk][CCP]])
\end{itemize}

@@

Discussion revolved around a few questions: what is the benefit of having a "community" for software sustainability, what practices and circumstances lead to having a community, how can funding help or hinder this process, and perhaps most importantly how can best practices be described and distilled into a document that can help new projects.

The benefits of having a community that were brought up were considered largely obvious. In addition to having advocates for the software, and a possible source of ``free'' contributions to the codebase, the community becomes a good source for requirements, feedback, and metrics. The software community can also act as "cheerleaders" who convince funders or other potential users to fund/use the software, and thus help sustain the software.

@@

Funding can help the process by encouraging both value to the community and high-quality user support. Only providing funding for the software development may create good software, but with less likelihood to have a real community. It was discussed that federal laboratories are a good incubator for software communities, and that a general facility like EarthCube is too dispersed to really make a community. Also, domain-specific groups within laboratories or universities might provide as an incubator for software communitites.

@@

The most important next steps is a ``Best Practice'' document, which would describe what successful projects with engaged communities look like, how to replicate this type of project, and look at end-of-life on a community project. Inputs to this document would include a software community survey of high functioning communities such as R Open Science, Python SciPy, OPeNDAP, and Unidata, with analysis of factors that feed into their success. Also references lik the "Art of Community" could be adapted and summarized for the science software community.

>>>>

Mozilla Science maintains a ``Working Open' Project Guide'' (http://mozillascience.github.io/leadership-training). From the introduction:
\begin{quote}
Working openly with contributors enables your
    community to learn how to build and collaborate together. This
    document is a guideline on how to work openly and involve others
    in your projects with Mozilla. We want to help you engage your
    community in a way that encourages contributors and builds other
    leaders.
 \end{quote}

Several books have been written about software communities:
\begin{itemize}
\item ``Art of Community'' by Jono Bacon. We could consider distilling this for scientific software.
\item Iain Larmour, from EPSRC inthe UK (EPSRC: https://www.epsrc.ac.uk/) (not sure who from mentioned UK Collaborative Computational Projects (CCP: http://ccp.ac.uk)
\end{itemize}

@@

Discussion revolved around a few questions: what is the benefit of having a ``community'' for software sustainability, what
practices and circumstances lead to having a community, how can funding help or hinder this process, and perhaps most
importantly how can best practices be described and distilled into a document that can help new projects.

The benefits of having a community that were brought up were considered largely obvious. In addition to having advocates for
the software, and a possible source of ``free'' contributions to the codebase, the community becomes a good source for
requirements, feedback, and metrics. The software community can also act as ``cheerleaders'' who convince funders or other
potential users to fund/use the software, and thus help sustain the software.

@@

Funding can help the process by encouraging both value to the community and high-quality user support. Only providing funding
for the software development may create good software, but with less likelihood to have a real community. It was discussed
that federal laboratories are a good incubator for software communities, and that a general facility like EarthCube is too
dispersed to really make a community. Also, domain-specific groups within laboratories or universities might provide as an
incubator for software communities.

@@

The most important next steps is a ``Best Practice'' document, which would describe what successful projects with engaged
communities look like, how to replicate this type of project, and look at end-of-life on a community project. Inputs to this
document would include a software community survey of high functioning communities such as R Open Science, Python SciPy,
OPeNDAP, and Unidata, with analysis of factors that feed into their success. Also references lik the ``Art of Community'' could
be adapted and summarized for the science software community.
